---
title: "host-microbiome"
author: "John McGInniss"
date: '2022-03-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

I'm interested in host-microbe interactions in the lung and how they influence lung health. Recently, our group performed a study looking at outcomes after lung transplant in relation to the lung  microbiome and inflammatory markers. A challenge was how to compare these two datasets.

The microbiome data is from 16S amplicon sequencing and the inflammatory data is from a 41-plex Luminex assay. So for each sample we have both data. My initial analysis basically looked at both separately and how they related to health and disease post-transplant and then correlations between microbial taxa and the inflammatory mediators. 

Inspired by my teams' work on agent based modeling, I wanted to take a modeling approach to the data which I think moves this in a more mechanistic direction. 

First:
1. Fit models to data
2. Simulate models that have an increase, decrease, no change
3. Compare actual models to simulated -- which best aligns?


## Load Packages
```{r packages}
library(tidyverse)
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
library(rstanarm)
library(skimr)
```

## Data

```{r data}
# My data
microbiome_cytokine <- 
  read_tsv("data/microbiome_cytokine.tsv")

#DF with sample, top taxa and inflammatory markers

glimpse(microbiome_cytokine)

# Distribution of cytokines
microbiome_cytokine %>% 
  pivot_longer(cols = c(1:42), names_to = "variable", values_to = "value") %>% 
  group_by(variable) %>% 
  summarise(median = median(value), iqr25 = quantile(value, 0.25), iqr75 = quantile(value, 0.75)) %>% 
  ggplot(., aes(x=reorder(variable, -median), y=median)) + 
  geom_point() + 
  geom_linerange(aes(ymin=iqr25, ymax=iqr75)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Distribution of taxa
microbiome_cytokine %>% 
  pivot_longer(cols = c(43:58), names_to = "variable", values_to = "value") %>% 
  group_by(variable) %>% 
  summarise(median = median(value), iqr25 = quantile(value, 0.25), iqr75 = quantile(value, 0.75)) %>% 
  ggplot(., aes(x=reorder(variable, -median), y=median)) + 
  geom_point() + 
  geom_linerange(aes(ymin=iqr25, ymax=iqr75)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Correlation plot
cor_df <- cor(scale(microbiome_cytokine), method = "spearman")
corrplot::corrplot(cor_df)

```
Overall there are modest correlations between taxa and cytokines. Does a modeling approach confirm this?


## Bayesian modeling
```{r}
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1) # Cauchy distribution with thick tails

set.seed(123)

# specify the parsnip model
bayes_mod <-   
  linear_reg() %>% 
  set_engine("stan", 
             prior_intercept = prior_dist, 
             prior = prior_dist) 

# train the model
bayes_fit <- bayes_mod %>% 
  fit(TNFa ~ Prevotella, data = microbiome_cytokine)

print(bayes_fit, digits = 5)

tidy(bayes_fit, conf.int = TRUE)

# Plot output
bayes_plot_data <- 
  new_points %>% 
  bind_cols(predict(bayes_fit, new_data = new_points)) %>% 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "urchin size") + 
  ggtitle("Bayesian model with t(1) prior distribution")
```

## Simulation models
```{r}
lm_mod <- 
  linear_reg() %>% 
  set_engine("lm")

# Fit the model to the data
lm_fit <- 
  lm_mod %>% 
  fit(RANTES ~ Prevotella, data = microbiome_cytokine)
lm_fit
tidy(lm_fit) # wrapper to put model in tidy form

# Take a look at model pars graphically
tidy(lm_fit) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```


